---
output: github_document
---

# CoMET: A Compressed Bayesian Mixed-Effects Model for High-Dimensional Tensors

This repositary contains:

1. **R package BayesCoMET**
    - Implements the Compressed Bayesian Mixed-Effects Model for Tensors (CoMET).
    - Install this library with:
    ```{r, eval = FALSE}
    devtools::install_github("SreyaSarkar21/BayesCoMET")
    library(BayesCoMET)
    ```
    
2. **Others Folder**
    - contains the codes implementing the competing methods.
    - `sampler_Oracle.R` contains the code for oracle, the Bayesian oracle benchmark of CoMET, where the true random-effects covariance structure is known.
    - `fanli2012.R` contains the code for implementing the penalized quasi-likelihood method for fixed effects selection by [1].
    - `licaili2021.R` contains the code implementing the penalized quasi-likelihood estimation and inference procedures for fixed effects selection by [2], using their published supplementary code as a reference.
    - `gee_cv.m`, `gee_equicorr_predict.m`, `gee_run_DEAM.m` are the MATLAB codes using the SparseReg MATLAB library for implementing the GEE approach [3]. 


## Example Implementation of CoMET

```{r MargPred, message=FALSE, warning=FALSE}
library(BayesCoMET)
pdims <- qdims <- c(32, 32)
n_train <- 100; n_test <- 50; m <- 3 ## 3 observations per subject

library(BayesCoMET)

## Generating a  CP-structured fixed-effect coefficient
cp_decomp_B <- BayesCoMET:::generate_cpB_with_sparse_factors(pdims = c(32, 32), K = 4,
                                                             sparsity = 0.25, seed = 123)
trueB <- cp_decomp_B$B

## Equicorrelation covariance matrices per tensor-mode
equicorr_mat <- function(q, rho) {
    if(rho < -1/(q-1) | rho > 1) {
        stop(sprintf("For q = %d, rho must be in [-1/(q-1), 1] for postive semi-definiteness.", q))
    }
    Sigma <- matrix(rho, q, q)
    diag(Sigma) <- 1
    Sigma
}

Sigma1 <- Sigma2 <- equicorr_mat(q = 32, rho = 0.5)
L1 <- chol(Sigma1); L2 <- chol(Sigma2)

## Simulating data with 150 subjects each with cluster size m
dat <- BayesCoMET:::simdata(pdims = c(32, 32), qdims = c(32, 32), n = 150, m = 3,
                            errVar = 0.1, B = trueB, L_list = list(L1 = L1, L2 = L2),
                            xcov_var = 1, zcov_var = 1, myseed = 1)

## Taking first 100 subjects for training and the rest for validation
y_train <- dat$yijs[1:(n_train * m)]; y_test <- dat$yijs[(n_train * m + 1):length(dat$yijs)]
xlist_train <- dat$Xijlist[1:(n_train * m)]; xlist_test <- dat$Xijlist[(n_train * m + 1):length(dat$yijs)]
zlist_train <- dat$Zijlist[1:(n_train * m)]; zlist_test <- dat$Zijlist[(n_train * m + 1):length(dat$yijs)]
nmodes <- length(pdims)

## Implement CoMET model Gibbs sampler
kdims <- rep(3, nmodes)
set.seed(12345)
R_list <- lapply(1:nmodes, function(d) {matrix(rnorm(kdims[d] * qdims[d], 0, sd = sqrt(1/kdims[d])), kdims[d], qdims[d])})
S_list <- lapply(1:nmodes, function(d) {matrix(rnorm(kdims[d] * qdims[d], 0, sd = sqrt(1/kdims[d])), kdims[d], qdims[d])})

res <- BayesCoMET::comet(y = y_train, xlist = xlist_train, zlist = zlist_train,
                         mis = rep(n_train, m),
                         K = 6, kdims = kdims, a0 = 0.01, b0 = 0.01, gammaVar0 = c(1, 1),
                         R_list = R_list, S_list = S_list,
                         niter = 1100, nburn = 100, nthin = 1, store_ranef = TRUE)

betaPostMed <- apply(res$betaSamp, 2, median)
vecB_true <- as.vector(trueB)
```

```{r visualize_betaSamps, echo = FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=2.5}
### beta_73 ###
# op <- par(no.readonly = TRUE)
# 
# layout(matrix(c(1, 2, 3, 3), nrow = 2, byrow = TRUE), heights = c(1, 0.25))
op <- par(mfrow = c(1, 2))
hist(res$betaSamp[, 73], freq = FALSE, col = "lightgrey", border = "black",
     main = "", xlab = expression(beta[73]), ylab = "",
     cex.lab = 0.85)
abline(v = quantile(res$betaSamp[, 73], c(0.025, 0.975)), col = "magenta", lty = 2, lwd = 2)
abline(v = quantile(res$betaSamp[, 73], 0.5), col = "blue", lty = 3, lwd = 2)
abline(v = vecB_true[73], col = "red", lty = 1, lwd = 2)

hist(res$betaSamp[, 1], freq = FALSE, col = "lightgrey", border = "black",
     main = "", xlab = expression(beta[1]), ylab = "", yaxt = "n",
     cex.lab = 0.85)
abline(v = quantile(res$betaSamp[, 1], c(0.025, 0.975)), col = "magenta", lty = 2, lwd = 2)
abline(v = quantile(res$betaSamp[, 1], 0.5), col = "blue", lty = 3, lwd = 2)
abline(v = vecB_true[1], col = "red", lty = 1, lwd = 2)
par(op)

op <- par(mfrow = c(1,1))
plot.new()
legend("center",
       legend = c("95% credible interval",
                  "Posterior median",
                  "True value"),
       col = c("magenta", "blue", "red"),
       lty = c(2, 3, 1),
       lwd = 2,
       horiz = TRUE,
       bty = "n")
par(op)
```


## Acknowledgement

Sreya Sarkar was supported by the National Science Foundation grant DMS-1854667 for this work.


## Citation

If you use *BayesCoMET* in your work, please cite:
Sarkar, S., Khare, K., & Srivastava, S. (2026). **CoMET: A Compressed Bayesian Mixed-Effects Model for High-Dimensional Tensors.** *arXiv.* [https://arxiv.org/pdf/2602.19236](https://arxiv.org/pdf/2602.19236)


## Other References

[1] Fan, Y., & Li, R. (2012). **Variable Selection in Linear Mixed Effects Models.** *The Annals of Statistics*, Vol. 40, No. 4, pp. 2043–2068. DOI: [10.1214/12-AOS1028](https://doi.org/10.1214/12-AOS1028).

[2] Li, S., Cai T. T., & Li, H. (2022). **Inference for High-Dimensional Linear Mixed-Effects Models: A Quasi-Likelihood Approach.** *Journal of the American Statistical Association*, 117(540), 1835–1846. DOI: [10.1080/01621459.2021.1888740](https://doi.org/10.1080/01621459.2021.1888740).

[3] Zhang X., Li, L., Zhou, H., Shen, D., et al. (2019). **Tensor generalized estimating equations for longitudinal imaging analysis.** *Statistica Sinica*, 29(4), 1977-2005. DOI: [10.5705/ss.202017.0153](https://doi.org/10.5705/ss.202017.0153).


